---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "80%"
)
```

# intervalpsych

<!-- badges: start -->
<!-- badges: end -->

The goal of intervalpsych is to help you with the analysis of interval responses.

## Installation

You can install the latest version of intervalpsych from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("matthiaskloft/intervalpsych")
```


# Fit the Interval Truth Model to the Verbal Quantifiers Dataset

First, we load the verbal quantifiers dataset:
```{r}
library(tidyverse)
library(intervalpsych)


data(quantifiers)

quantifiers <- quantifiers |>
  # exclude control items
  dplyr::filter(!name_en %in% c("always", "never", "fifty-fifty chance")) |>
  # sample 100 respondents
  dplyr::filter(id_person %in% sample(
    size = 100,
    replace = FALSE,
    unique(quantifiers$id_person)
  )) |>
  # exclude missing values
  dplyr::filter(!is.na(x_L) & !is.na(x_U)) |>
  # recompute IDs
  mutate(
    id_person = factor(id_person) |> as.numeric(),
    id_item = factor(id_item) |> as.numeric()
  )

head(quantifiers)
```


First, we need to convert the interval responses to the simplex format:

```{r}
quantifiers <- cbind(
  quantifiers, 
  itvl_to_splx(quantifiers[,c("x_L","x_U")], min = quantifiers$scale_min, max = quantifiers$scale_max))

head(quantifiers[,9:13])
```


Let's check if we can apply the Isometric Log-Ratio transformation:
```{r}
try(ilr(quantifiers[,c("x_1","x_2","x_3")]))
```

It seems we have components in our simplex data that are zero. So we first have
to deal with these zero components. We can do this by adding a padding constant:

```{r}
quantifiers[, c("x_1", "x_2", "x_3")] <- 
  remove_zeros(quantifiers[, c("x_1", "x_2", "x_3")], padding = 0.01)

head(quantifiers[,9:13])
```

```{r}
fit <-
  fit_itm(
    df_simplex = quantifiers[, c("x_1", "x_2", "x_3")],
    id_person = quantifiers$id_person,
    id_item = quantifiers$id_item,
    item_labels = quantifiers |> 
      distinct(id_item, name_en) |> 
      pull(name_en),
    n_chains = 2,
    n_cores = 2,
    iter_sampling = 300,
    iter_warmup = 300,
    adapt_delta = .95
  )
```

Now we can extract the estimated cosensus intervals from the fit object:
```{r}
consensus <- extract_consensus(fit)
```

```{r}
plot_intervals(
  df_interval_bounds = consensus$summary[,c("T_L_median", "T_U_median")],
  item_labels = rownames(consensus$summary))
```












